\documentclass[12pt]{article}
\usepackage{assignment-style}
\usepackage{float}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{xcolor}

\lstdefinelanguage{Cypher}{
  keywords={MATCH, WHERE, RETURN, AND, OR, NOT},
  sensitive=true,
  morecomment=[l]{//},
  morestring=[b]"
}

\lstset{
  language=Cypher,
  basicstyle=\ttfamily,
  keywordstyle=\color{blue}\bfseries,
  columns=fullflexible,
  keepspaces=true,
  stringstyle=\color{red},
  showstringspaces=false,
  breaklines=true
}


\graphicspath{ {./images/} }

\papercode{COMPSCI752}
\title{Assignment 2 report}

\author{Romili \textsc{Townsend}}
\studentid{rtow454}

\reportdate{\today}

\begin{document}

\maketitle

1. Write down the formal property graph model that corresponds to the
property graph G in Figure 1. This means for $G = (V, E, \eta, \lambda, \nu)$
specify the sets / (partial) functions $V$, $E$, $\eta$, $\lambda$ and $\nu$.
[10 marks]

\begin{equation*}
\begin{aligned}
    V &: 1, 2, 3 \\
    E &: 4, 5, 6 \\
    \eta &: 4 \to (1,2), 5 \to (2,3), 6 \to (1,3) \\\
    \lambda &: 1 \to \{\text{Client}\}, 2 \to \{\text{Client}\}, 3 \to \{\text{Client}\}, \\
    &\quad 4 \to \{\text{Transaction}\}, 5 \to \{\text{Transaction}\}, 6 \to \{\text{Transaction}\} \\
    \nu &: \\
    &\text{(1, name)} \to \text{'Aramis'}, \\
    &\text{(2, name)} \to \text{'Athos'}, \\
    &\text{(3, name)} \to \text{'Porthos'}, \\
    &\text{(4, date)} \to \text{29:05:25}, \\
    &\text{(4, amount)} \to \text{20k}, \\
    &\text{(5, date)} \to \text{30:05:24}, \\
    &\text{(5, amount)} \to \text{25k}, \\
    &\text{(6, date)} \to \text{01:06:24}, \\
    &\text{(6, amount)} \to \text{15k} \\
\end{aligned}
\end{equation*}

2. For each of the following English language queries, write it down
in Regular Property Graph Logic. You do not have to evaluate the
queries on the graph G in Figure 1.

\begin{enumerate}[label=(\alph*)]
    % (a) %
    \item Return pairs of clients, c1 and c2 where client c1 transferred more
    than 5k to client c2. [2 marks]
    
    \begin{equation*}
    \begin{aligned}
        \text{result}(c_1, c_2) \leftarrow \text{:Transaction}(c_1, c_2) \text{ AS } t, \text{ } t.\text{amount} >5000. \\
    \end{aligned}
    \end{equation*}

    % (b) %
    \item Return pairs of clients in which client 1 initiates a transaction t
    to client 2 and client 2 initiates two transactions t1 and t2, both
    to client 1, where the total amount of t1 and t2 is the same as the
    amount of t, and t1 and t2 where initiated on different dates.
    [4 marks]

    \begin{equation*}
    \begin{aligned}
        \text{result}(c_1, c_2) \leftarrow \quad
        & \text{:Transaction}(c_1, c_2) \text{ AS }t,              \\
        & \text{:Transaction}(c_2, c_1) \text{ AS }t_1,            \\
        & \text{:Transaction}(c_2, c_1) \text{ AS }t_2,            \\
        & t_1 \neq t_2,                                            \\
        & t_1.\text{date} \neq t_2.\text{date},                    \\
        & t.\text{amount} = t_1.\text{amount} + t_2.\text{amount}. \\
    \end{aligned}
    \end{equation*}

    % (c) %
    \item A fraud ring of length $l$ has n clients where for $i = 1, . . . , l - 1$,
    client i initiates a transaction to client $i + 1$ with an amount of
    at least 50k, and client l initiates a transaction to client $1$ with
    an amount of at least 50k.
    Return client $1$ in any fraud ring of size $l$ where $l$ is an odd integer
    larger than $1$. [4 marks]

    \begin{equation*}
    \begin{aligned}
        \text{result}(c_1) \leftarrow \quad 
        & \text{:Transaction*}(c_1, c_l) \text{ AS } t, \text{ } t.\text{amount} \ge 50000, \\
        & \text{:Transaction}(c_l, c_1) \text{ AS } t', \text{ } t'.\text{amount} \ge 50000, \\
        & l > 1, \text{ } l \bmod 2 \neq 0.
    \end{aligned}
    \end{equation*}

\end{enumerate}

3. Consider again the three queries from question 2. Write down all three
queries from question 2 in Regular Property Graph Algebra. You do not
have to evaluate the queries on the graph G in Figure 1. [10 marks]


\begin{enumerate}[label=(\alph*)]
    % (a) %
    \item 
    
    \begin{equation*}
    \begin{aligned}
        \bowtie^{\Phi}_{\text{src, trg}}\sigma_{\text{amount} > 5000}\text{(:Transaction)}
    \end{aligned}
    \end{equation*}

    % (b) %
    \item

    \begin{equation*}
    \begin{aligned}
        \bowtie^{\text{\O}}_{\text{src, trg}}
        \Big(
            \bowtie^{\Phi}_{\text{src$_1$, trg$_1$}}\text{(:Transaction) }\cup
            \bowtie^{\Phi}_{\text{src$_2$, trg$_2$}}\text{(:Transaction) }\cup
            \bowtie^{\Phi}_{\text{src$_3$, trg$_3$}}\text{(:Transaction)}
        \Big) \\
        \noalign{where $\Phi$ is src$_2=$ src$_3$, trg$_1=$ src$_2$, trg$_2=$ src$_1$, trg$_3=$ src$_1$,}
        \noalign{edge$_1$.amount = edge$_2$.amount + edge$_3$.amount}
    \end{aligned}
    \end{equation*}

    % (c) %
    \item 

    % TODO: Complete 3c

    \begin{equation*}
    \begin{aligned}
        \bowtie^{\Phi}_{\text{src}}
    \end{aligned}
    \end{equation*}

\end{enumerate}

4. Write down all three queries from question 2 in Cypher. For this you
might want to create a sample graph dataset similar to the graph in
Figure 1 in Neo4j and test your answers. You do not have to evaluate
the queries on the graph G in Figure 1.

\begin{enumerate}[label=(\alph*)]
    % TODO: Complete all 3 of these equations

    % (a) %
    \item
    \begin{lstlisting}
    MATCH(c1:Client)-[t:Transaction]->(c2:Client)
    WHERE t.amount > 5000
    RETURN c1, c2
    \end{lstlisting}


    % (b) %
    \item
    \begin{lstlisting}
    MATCH (c2:Client)-[t1:Transaction]->(c1:Client),
    (c1)-[t:Transaction]->(c2)-[t2:Transaction]->(c1)
    WHERE t1 <> t2
        AND t1.date <> t2.date
        AND t.amount = t1.amount + t2.amount
    RETURN c1, c2
    \end{lstlisting}

    % (c) %
    \item
\end{enumerate}

\end{document} 